---
title: "Machine Learning :PROJECT"
author: "Nikhil Prakash"
---
#### Introduction
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. Our goal is to predict the manner in which they did the exercise.

```{r results="hide", package_options, include=FALSE}
library(caret)
library(rpart)
library(rpart.plot)
library(RColorBrewer)
library(rattle)
library(randomForest)
library(knitr)
```

#### Download and Load the data.
```{r}
set.seed(12345)
if(!file.exists("pml-training.csv")){
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", destfile = "pml-training.csv")}
trainData<- read.csv("pml-training.csv", sep=",", header=TRUE, na.strings = c("NA","","#DIV/0!"))
if(!file.exists("pml-testing.csv")){
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", destfile = "pml-testing.csv")}
testData<- read.csv("pml-testing.csv", sep=",", header=TRUE, na.strings = c("NA","","#DIV/0!"))
dim(trainData)
dim(testData)
```

#### Cleaning the data.

Remove NearZeroVariance variables
```{r}
nzv <- nearZeroVar(trainData, saveMetrics=TRUE)
trainData <- trainData[,nzv$nzv==FALSE]
nzv<- nearZeroVar(testData,saveMetrics=TRUE)
testData <- testData[,nzv$nzv==FALSE]
```

Partition the trainData into two trainData_TRAIN & trainData_TEST
```{r}
partition <- createDataPartition(trainData$classe, p=0.6, list=FALSE)
trainData_TRAIN <- trainData[partition, ]
trainData_TEST <- trainData[-partition, ]
dim(trainData_TRAIN)
dim(trainData_TEST)
```

Remove first column as it contain serial number and Clean variables with more than 60% NA.
```{r}
trainData_TRAIN <- trainData_TRAIN[c(-1)]
cleanV <- trainData_TRAIN
for(i in 1:length(trainData_TRAIN)) {
    if( sum( is.na( trainData_TRAIN[, i])) /nrow(trainData_TRAIN) >= .7){
        for(j in 1:length(cleanV)){
            if( length( grep(names(trainData_TRAIN[i]), names(cleanV)[j])) == 1){
                cleanV <- cleanV[ , -j]
            }   
        } 
    }
}
trainData_TRAIN <- cleanV
```

Restructure the trainData_TEST and testData.
```{r}
temp1 <- colnames(trainData_TRAIN)
temp2 <- colnames(trainData_TRAIN[, -58])
trainData_TEST <- trainData_TEST[temp1]
testData <- testData[temp2]
dim(trainData_TEST)
dim(testData)
```

Make all the data into same data type
```{r}
for (i in 1:length(testData) ) {
    for(j in 1:length(trainData_TRAIN)) {
        if( length( grep(names(trainData_TRAIN[i]), names(testData)[j]) ) == 1)  {
            class(testData[j]) <- class(trainData_TRAIN[i])
        }      
    }      
}
testData <- rbind(trainData_TRAIN[2, -58] , testData)
testData <- testData[-1,]
```

Prediction using Decision Trees
```{r}
set.seed(12345)
DecisionTree <- rpart(classe ~ ., data=trainData_TRAIN, method="class")
fancyRpartPlot(DecisionTree)
Pred_DecisionTree <- predict(DecisionTree, trainData_TRAIN, type = "class")
DTree <- confusionMatrix(Pred_DecisionTree, trainData_TRAIN$classe)
DTree
plot(DTree$table, col = DTree$byClass, main = paste("Decision Tree Confusion Matrix: Accuracy =", round(DTree$overall['Accuracy'], 4)))
```

Prediction using Random Forests
```{r}
set.seed(12345)
RandomForest <- randomForest(classe ~ ., data=trainData_TRAIN)
pred_RandomForest <- predict(RandomForest, trainData_TRAIN, type = "class")
RForest <- confusionMatrix(pred_RandomForest, trainData_TRAIN$classe)
RForest
plot(RandomForest)
plot(RForest$table, col = RForest$byClass, main = paste("Random Forest Confusion Matrix: Accuracy =", round(RForest$overall['Accuracy'], 4)))
```

Prediction using Generalized Boosted Regression
```{r}
set.seed(12345)
fitControl <- trainControl(method = "repeatedcv", number = 5, repeats = 1)
GenBoost <- train(classe ~ ., data=trainData_TRAIN, method = "gbm", trControl = fitControl, verbose = FALSE)
GenBoost_FM <- GenBoost$finalModel
Pred_GenBoost <- predict(GenBoost, newdata=trainData_TRAIN)
Acc_GenBoost <- confusionMatrix(Pred_GenBoost, trainData_TRAIN$classe)
Acc_GenBoost
plot(GenBoost, ylim=c(0.9, 1))
```

#### Conclusion
If we compare the above predection based on Accuracy for the Random forest was more accurate then the Decision Trees or GBM.
The expected out-of-sample error is 0 percent.

#### Using Random Forest Predict result on the test data.
```{r}
FinalPred <- predict(RandomForest, testData, type = "class")
FinalPred
```
